### 2008-03-13 trunk/Papers/dl-gre-08/bisim.tex
More text and added picture.

Formulas of both $\alc$ and $\el$ are interpreted in relational first-order models
$\gM = (\Delta^\gM,|\cdot|^\gM)$ where $W$ is an arbitrary non-empy set, and $|\cdot|^\gM$ is an interpretation
function such that:
$$
\begin{array}{ccl}
|p|^\gM & \subseteq & \Delta^\gM  \mbox{ for $p \in \prop$}\\
|R|^\gM & \subseteq & \Delta\times W  \mbox{ for $R \in \rel$}\\
|\neg \varphi|^\gM & = & W \backslash |\varphi|^\gM\\
|\varphi \wedge \varphi'|^\gM & = & |\varphi|^\gM \cap |\varphi'|^\gM\\
|\exists R.\varphi|^\gM & = & \{o \mid \mbox{for some } o' ((o,o') \in |R|^\gM)\\
& & \mbox{ and } o' \in |\varphi|^\gM \}.\\
\end{array}
$$
\end{definition}

We can now exactly characterize the problem of when a formula distinguishing two
elements in a given model exists. We will do so using the notion of \emph{bisimulaiton}.

\begin{definition}
Given a model $\gM = (\Delta^\gM,)$
\end{definition}

>>>>

Formulas of both $\alc$ and $\el$ are interpreted in relational first-order models
$\gM = (\Delta^\gM,|\cdot|^\gM)$ where $W$ is an arbitrary non-empy set, and $|\cdot|^\gM$ is an interpretation
function such that:
$$
\begin{array}{ccl}
|p|^\gM & \subseteq & \Delta^\gM  \mbox{ for $p \in \prop$}\\
|R|^\gM & \subseteq & \Delta\times W  \mbox{ for $R \in \rel$}\\
|\neg \varphi|^\gM & = & W \backslash |\varphi|^\gM\\
|\varphi \wedge \varphi'|^\gM & = & |\varphi|^\gM \cap |\varphi'|^\gM\\
|\exists R.\varphi|^\gM & = & \{o \mid \mbox{for some } o' ((o,o') \in |R|^\gM)\\
& & \mbox{ and } o' \in |\varphi|^\gM \}.\\
\end{array}
$$
Given a model $\gM$ and an object $o$ in the domain of $\gM$, let
$\propm(o) = \{p \in \prop \mid o \in |p|^\gM\}$.
\end{definition}

We can now exactly characterize the problem of when a formula distinguishing two
elements in a given model exists. We will do so using the notion of \emph{bisimulation}.

\begin{definition}
Given a model $\gM = (\Delta^\gM,|\cdot|^\gM)$, two elements $o$ and $o'$ of $\Delta^\gM$ are $\alc$-bisimilar if it is possible to find a relation $Z \subseteq \Delta^\gM \times \Delta^\gM$ such that
\begin{enumerate}
\item $(o,o') \in Z$.
\item if $(e_1, e_2) \in Z$ then $\propm(e_1) = \propm(e_2)$.
\item if $(e_1,e_2) \in Z$ and for some $R \in \rel$ there is $e_1'$ such that
$(e_1,e_1') \in |R|^\gM$ then there is $e_2'$ such that $(e_2,e_2') \in |R|^\gM$ and
$(e_1',e_2') \in Z$.
\item if $(e_1,e_2) \in Z$ and for some $R \in \rel$ there is $e_2'$ such that
$(e_2,e_2') \in |R|^\gM$ then there is $e_1'$ such that $(e_1,e_1') \in |R|^\gM$ and
$(e_1',e_2') \in Z$.
\end{enumerate}
\end{definition}

\begin{theorem}[\cite{modallogic}]
Given a model $\gM$, if  $o$ and $o'$ are two $\alc$-bisimilar elements in $\gM$
then there is no $\alc$-formula $\varphi$ such that $o \in |\varphi|^\gM$ and
$o' \not \in |\varphi|^\gM$.
\end{theorem}

In other words, $\alc$-bisimilar elements are indistinguishable in the $\alc$ langauge.
By defining a suitable notion of $\el$-bisimulation, the same result can be proved for
the $\el$ language.
